---
title: "Lab 4_v2"
author: "Karisa Ke"
format: html
editor: visual
embed-resources: true
---

## 1. Read in the data

```{r}
library(tidyverse)
library(data.table)
```

```{r}
if (!file.exists("met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )
met <- data.table::fread("met_all.gz")
```

## 2. Prepare the data

-   Remove temperatures less than -17C

```{r}
met <- met[temp>-17]
```

-   Make sure there are no missing data in the key variables coded as 9999, 999, etc

```{r}
met[met$elev==9999.0] <- NA
```

```{r}
summary(met$dew.point)
```

-   Generate a date variable using the functions `as.Date()`

```{r}
met$date <- as.Date(paste(met$year, met$month, met$day, sep = "-"))
```

```{r}
met$week = week(met$date)
```

```{r}
met =filter(met, week < 33 & day < 8)
```

```{r}
met
```

```{r}
met_avg <- met[,.(temp=mean(temp,na.rm=TRUE), rh=mean(rh,na.rm=TRUE), wind.sp=mean(wind.sp,na.rm=TRUE), vis.dist=mean(vis.dist,na.rm=TRUE), dew.point = mean(dew.point, na.rm=TRUE), lat=mean(lat), lon=mean(lon), elev=mean(elev,na.rm=TRUE)), by="USAFID"]
```

```{r}
met_avg$region <- ifelse(met_avg$lon > -98 & met_avg$lat >39.71, "NE", ifelse(met_avg$lon > -98 & met_avg$lat < 39.71, "SE", ifelse(met_avg$lon < -98 & met_avg$lat >39.71, "NW", "SW")))
```

```{r}
met_avg$elev_cat <- ifelse(met_avg$elev> 252, "high", "low")
```

```{r}
table(met_avg$region)
```

## 3. **Use `geom_violin` to examine the wind speed and dew point by region**

```{r}
met_avg %>%
  filter(!(region %in% NA)) %>%
ggplot()+
  geom_violin(mapping = aes(y=wind.sp, x=1)) +
  facet_wrap(~region, nrow=2)
```

```{r}
met_avg %>%
  filter(!(region %in% NA)) %>%
ggplot()+
  geom_violin(mapping = aes(y=dew.point, x=1)) +
  facet_wrap(~region, nrow=2)
```

## 4. **Use `geom_jitter` with `stat_smooth` to examine the association between dew point and wind speed by region**

```{r}
met_avg %>%
filter(!(region %in% NA)) %>%
  ggplot(mapping = aes(x=dew.point, y=wind.sp, color=region))+
  geom_jitter() + 
  stat_smooth(method=lm)
```

## 5. **Use `geom_bar` to create barplots of the weather stations by elevation category colored by region**

```{r}
met_avg %>%
filter(!(region %in% NA)) %>%
  ggplot()+
  geom_bar(mapping=aes(x=elev_cat,fill=region), position = "dodge")+
  scale_fill_brewer(palette = "PuOr")+
  labs(title="Number of weather stations by elevation category and region", x="Elevation Category", y= "Count")+
  theme_bw()
```

## 6. **Use `stat_summary` to examine mean dew point and wind speed by region with standard deviation error bars**

```{r}
met_avg %>%
filter(!(region %in% NA)) %>%
  ggplot(mapping=aes(x=region, y=dew.point)) +
  stat_summary(fun.data="mean_sdl", geom="errorbar") +
  stat_summary(fun.data="mean_sdl")Ye1
```

```{r}
met_avg %>%
filter(!(region %in% NA)) %>%
  ggplot(mapping=aes(x=region, y=wind.sp)) +
  stat_summary(fun.data="mean_sdl", geom="errorbar") +
  stat_summary(fun.data="mean_sdl")
```

```{r}

```

## 7. **Make a map showing the spatial trend in relative humidity in the US**

```{r}
library(leaflet)
met_avg2<-met_avg[!is.na(rh)]

tops <- met_avg2[rank(-rh) <= 10]

rh_pal = colorNumeric(c('red','yellow','blue'), domain=met_avg2$rh)
leaflet(met_avg2) %>%
  addProviderTiles('OpenStreetMap') %>%
  addCircles(lat=~lat, lng=~lon, color=~rh_pal(rh), label=~paste0(round(rh,2), ' rh'), opacity=1,fillOpacity=1, radius=500) %>%
  addMarkers(lat=~lat, lng=~lon, label=~paste0(round(rh,2), ' rh'), data = tops) %>%
  addLegend('bottomleft',pal=rh_pal, values=met_avg2$rh, title="Relative Humidity", opacity=1)
```

## 8. Use a ggplot extension

```{r}
install.packages("ggforce")
```

```{r}
library(ggforce)
ggplot(met_avg, aes(wind.sp, dew.point, colour = region)) +
  geom_point() +
  facet_zoom(x = Species == "versicolor")
```
